Project Diary

### 2014-05-10 thesis/cross_correlation_gaussian.tex
Revision following Sven's advice.

\chapter{Correlation Matrix of Gaussian Returns}\label{chp:Gaussian}
In this chapter we present some analytical and numerical results about
the cross-correlation matrix (cf. \S
\ref{sec:FundamentalConcepts}), in the special case where the involved
asset returns are assumed to be Gaussian distributed. This assumption
is of course an over-simplification, but nevertheless lends some
insight into the problem.

Our primary interest is in the influence of autocorrelations on the
cross-correlation matrix. \S \ref{sec:GCC-analytical} discusses the
distribution of the matrix entries and \S \ref{sec:GCC-numerical} the
distribution of the eigenvalues.

\section{Distribution of the Matrix Elements}\label{sec:GCC-analytical}
If the returns follow a zero-mean Gaussian distribution, 
i.e. $\vec{r}_t \sim N(0, \Sigma)$, and are not
auto-correlated, the L\'evy index $\alpha$ as appears in equation
\ref{eq:cross-correlation} is 2, and $RR'$ is a Wishart matrix with
the following distribution law \cite{Anderson2003}:
\begin{equation}
  \label{eq:WishartPDF}
  f_{RR'}(X) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
      \Sigma^{-1}X \right)
    \over
    2^{NT/2}\pi^{N(N-1)/4}|\det \Sigma|^{T/2}
    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
  }
\end{equation}
where $\Sigma$ is the true covariance matrix of the returns.

@@

As before we consider the AR(1) process (see
\S\ref{sec:FundamentalConcepts}):
\begin{equation*}
  \vec{r}_t = \phi \vec{r}_{t-1} + \vec{a}_{t}
\end{equation*}
where $\vec{a}_t \sim N(0, I)$, i.e. the elements of $\vec{a}_t$ are
independent Gaussian random variable with zero mean and unit
variance.

>>>>

\chapter{Correlation Matrix of Gaussian Returns}\label{chp:Gaussian}
In this chapter we present some results about the cross-correlation
matrix of Gaussian returns. This assumption is of course an
over-simplification, but nevertheless lends some insight into the
problem.

Our primary interest is in the influence of autocorrelations on the
cross-correlation matrix. \S\ref{sec:GCC-analytical} and
\S\ref{sec:GCC-numerical} discuss the distribution of the matrix
elements and the eigenvalues, respectively.

\section{Distribution of the Matrix Elements}\label{sec:GCC-analytical}
If the returns follow a zero-mean Gaussian distribution, 
i.e. $\vec{r}_t \sim N(0, \Sigma)$, and are not
auto-correlated, then $RR'$ is a Wishart matrix with
the following distribution law \cite{Anderson2003}:
\begin{equation}
  \label{eq:WishartPDF}
  f_{RR'}(X) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
      \Sigma^{-1}X \right)
    \over
    2^{NT/2}\pi^{N(N-1)/4}|\det \Sigma|^{T/2}
    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
  }
\end{equation}
where $\Sigma$ is the true covariance matrix of the returns.

@@

As before we consider the AR(1) process:
\begin{equation*}
  \vec{r}_t = \phi \vec{r}_{t-1} + \vec{a}_{t}
\end{equation*}
where $\vec{a}_t \sim N(0, I)$, i.e. the elements of $\vec{a}_t$ are
independent Gaussian random variable with zero mean and unit
variance.
